---
title: "Maurum Analysis"
author: Saydie Sago, adapted from "Heather Deel" 2025-05-14 and Lyanna Kessler 2023-09-22  
date: "2025-08-20
output:html_document: default
---

# This script was modeled by Heather Deel's script for the QIIME2 microbiome workshop on 5/14/25 and 6/10-12/25 and have taken aspects of Lyanna Kessler's script from the Desmond et. al., 2025 publication to generate the M.aurum microbiome analysis. I started working on the OnDemand RC Colorado (version 3.1.14) Rstudio server on "Interactive Session" tab. The latest RStudio version OnDemand RC uses is 2024.04.2 R 4.4.1, and I selected 1 Core, 12 hours settings. Note as a reminder for CU Boulder students, you need your Identikey and password to login into OnDemand RC. If you don't want to use OnDemand RC, you can run this Rmd thorugh Rstudio on your desktop. 

#I had to switch over Rstudio on my desktop on 8/27/25 later during my analysis. OnDemand RC cannot download all microbiome analysis packages (e.g. microbiomeMarker and devtools) needed for complex analysis and Lowry lab approved. OnDemand RC has restrictions to add new directories to your RC account. Take home message, OnDemand RC uses basic microbiome packages, but if you are doing more an extensive analysis that need specific and special packages on GitHub, then work on your desktop :)  

# References
# Hex codes for M.aurum publication: BBS/HC = #B3E1F0, BBS/IS = #148CBA, Maurum/HC = #98E0B1, Maurum/IS = #21A951

################################################################################
# Installing and loading packages
################################################################################

## Install special packages for microbiome analysis and generating clean figures if it's not in your libary and the first time going through this script. Note it can take awhile.
```{r}
# Some packages are going to be installed via github/biocmanager or devtools 
remotes::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis", force = T)
remotes::install_github("phytomosaic/ecole")
remotes::install_github("joey711/phyloseq")
remotes::install_github("KarstensLab/microshades", dependencies = TRUE)
remotes::install_github("yiluheihei/microbiomeMarker")
remotes::install_github("taryue/GMDecomp")
remotes::install_github("phytomosaic/ecole")
BiocManager::install("philr")
BiocManager::install("ggtree")
BiocManager::install("microbiomeMarker") #cannot install because OnDemand RC cannot make a new R directory! 
BiocManager::install("ANCOMBC")
BiocManager::install("EnhancedVolcano")
devtools::install_github("jbisanz/qiime2R")
devtools::install_github("kylebittinger/usedist")
devtools::install_github("microbiome/microbiome")

```

## Upload packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
skoden_packages <- c("tidyverse",
                   "data.table",
                   "ggplot2",
                   "phyloseq",
                   "patchwork",
                   "microshades",
                   "cowplot",
                   "magrittr",
                   "qiime2R",
                   "vegan",
                   "microbiomeMarker",
                   "BiocManager",
                   "ANCOMBC",
                   "GMDecomp",
                   "ape",
                   "usedist",
                   "lme4",
                   "ellipse", 
                   "ggrepel",
                   "EnhancedVolcano",
                   "biomformat",
                   "ggdist",
                   "pairwiseAdonis",
                   "seqinr",
                   "plotrix",
                   "ggthemes",
                   "glmnet",
                   "ggtree",
                   "philr",
                   "ggpubr", 
                    "tibble", 
                    "stringr", 
                    "forcats", 
                    "microbiome",
                    "svglite", 
                    "systemfonts", 
                    "Rmisc",
                    "openxlsx")

lapply(skoden_packages, library, character.only=TRUE)
#install.packages("car") # needed to update car package on my computer, if you do, say no to "install from sources the packages which need compilation"
#"ecole" package does not exist and built for another older version of Rstudio 
```
################################################################################
# Alpha diversity
################################################################################
## Import alpha diversity data from QIIME2 output
```{r}
# import vectors exported from QIIME2
faith_pd <- read_tsv("output_for_R/faith_pd_vector.tsv")
evenness <- read_tsv("output_for_R/evenness_vector.tsv")
shannon <- read_tsv("output_for_R/shannon_vector.tsv")
obs_feats <- read_tsv("output_for_R/observed_features_vector.tsv")

# change sample ID column names
colnames(faith_pd) <- c("SampleID","faith_pd")
colnames(evenness) <- c("SampleID","pielou_evenness")
colnames(shannon) <- c("SampleID","shannon_entropy")
colnames(obs_feats) <- c("SampleID","observed_features")

# merge vectors by sample ID
alpha_div <- faith_pd %>% 
  left_join(evenness, by = "SampleID") %>% 
  left_join(shannon, by = "SampleID") %>% 
  left_join(obs_feats, by = "SampleID")

# what are we missing from this data frame? metadata!
metadata <- read_tsv("Maurum_only_metadata.tsv")

alpha_div_metadata <- alpha_div %>% 
  left_join(metadata, by = "SampleID")

```
## Statistical Analysis: 2x2 ANOVA for Faiths, Evenness, Entropy, OTUS
```{r}
alpha_div_metadata$treatment <- as.factor(alpha_div_metadata$Treatment)
alpha_div_metadata$condition <- as.factor(alpha_div_metadata$Condition)
alpha_div_metadata$experimentalgroup <- as.factor(alpha_div_metadata$experimental_group)

# Run two-way ANOVA for Faith's PD
aov1 <- aov(faith_pd ~ treatment * condition, data = alpha_div_metadata)

# Extract the summary table
anova_table1 <- summary(aov1)[[1]]

# Pull the p-values
p_treatment_faith       <- anova_table1["treatment", "Pr(>F)"]
p_condition_faith  <- anova_table1["condition", "Pr(>F)"]
p_interaction_faith <- anova_table1["treatment:condition", "Pr(>F)"]

# Make a nice label string
p_label_faith <- paste0("treatment p = ", signif(p_treatment_faith, 3),
                  "\ncondition p = ", signif(p_condition_faith, 3),
                  "\ninteraction p = ", signif(p_interaction_faith, 3))

# Run two-way ANOVA for Pielou Evenness 
aov2 <- aov(pielou_evenness ~ treatment * condition, data = alpha_div_metadata)

# Extract the summary table
anova_table2 <- summary(aov2)[[1]]

# Pull the p-values
p_treatment_evenness      <- anova_table2["treatment", "Pr(>F)"]
p_condition_evenness  <- anova_table2["condition", "Pr(>F)"]
p_interaction_evenness <- anova_table2["treatment:condition", "Pr(>F)"]

# Make a nice label string
p_label_evenness <- paste0("treatment p = ", signif(p_treatment_evenness, 3),
                  "\ncondition p = ", signif(p_condition_evenness, 3),
                  "\ninteraction p = ", signif(p_interaction_evenness, 3))


# Run two-way ANOVA for Shannon Diversity  
aov3 <- aov(shannon_entropy ~ treatment * condition, data = alpha_div_metadata)

# Extract the summary table
anova_table3 <- summary(aov3)[[1]]

# Pull the p-values
p_treatment_shannon       <- anova_table3["treatment", "Pr(>F)"]
p_condition_shannon  <- anova_table3["condition", "Pr(>F)"]
p_interaction_shannon <- anova_table3["treatment:condition", "Pr(>F)"]

# Make a nice label string
p_label_shannon <- paste0("treatment p = ", signif(p_treatment_shannon, 3),
                  "\ncondition p = ", signif(p_condition_shannon, 3),
                  "\ninteraction p = ", signif(p_interaction_shannon, 3))

# Run two-way ANOVA for OTUS
aov4 <- aov(observed_features ~ treatment * condition, data = alpha_div_metadata)

# Extract the summary table
anova_table4 <- summary(aov4)[[1]]

# Pull the p-values
p_treatment_otus       <- anova_table4["treatment", "Pr(>F)"]
p_condition_otus  <- anova_table4["condition", "Pr(>F)"]
p_interaction_otus <- anova_table4["treatment:condition", "Pr(>F)"]

# Make a nice label string
p_label_otus <- paste0("treatment p = ", signif(p_treatment_otus, 3),
                  "\ncondition p = ", signif(p_condition_otus, 3),
                  "\ninteraction p = ", signif(p_interaction_otus, 3))

summary_faithpd <- alpha_div_metadata %>%
  group_by(experimental_group) %>%
  summarise(
    n = sum(!is.na(faith_pd)),
    mean = mean(faith_pd, na.rm = TRUE),
    sd = sd(faith_pd, na.rm = TRUE),
    se = sd / sqrt(n),
    .groups = "drop"
  )
summary_faithpd

summary_evenness <- alpha_div_metadata %>%
  group_by(experimental_group) %>%
  summarise(
    n = sum(!is.na(pielou_evenness)),
    mean = mean(pielou_evenness, na.rm = TRUE),
    sd = sd(pielou_evenness, na.rm = TRUE),
    se = sd / sqrt(n),
    .groups = "drop"
  )
summary_evenness

summary_shannon <- alpha_div_metadata %>%
  group_by(experimental_group) %>%
  summarise(
    n = sum(!is.na(shannon_entropy)),
    mean = mean(shannon_entropy, na.rm = TRUE),
    sd = sd(shannon_entropy, na.rm = TRUE),
    se = sd / sqrt(n),
    .groups = "drop"
  )
summary_shannon

summary_otus <- alpha_div_metadata %>%
  group_by(experimental_group) %>%
  summarise(
    n = sum(!is.na(observed_features)),
    mean = mean(observed_features, na.rm = TRUE),
    sd = sd(observed_features, na.rm = TRUE),
    se = sd / sqrt(n),
    .groups = "drop"
  )
summary_otus
```
## Make figure- Treatment
```{r}
# Wilcoxon test:	Compare two groups (non-parametric)
# Kruskal-Wallis: Compare multiple groups (non-parametric)
# Two-Way ANOVA: continuous data and there is only one collection point(parametric)

# rebuild safely from Treatment + Condition
alpha_div_metadata$experimental_group <- paste0(
  tolower(as.character(alpha_div_metadata$Treatment)),
  ".",
  as.character(alpha_div_metadata$Condition)
)

# order + relabel groups
alpha_div_metadata$experimental_group <- factor(
  alpha_div_metadata$experimental_group,
  levels = c("bbs.HC", "bbs.IS", "maurum.HC", "maurum.IS"),
  labels = c("BBS/HC", "BBS/IS", "M. aurum/HC", "M. aurum/IS")
)

group_colors <- c(
  "BBS/HC" = "#B3E1F0",
  "BBS/IS" = "#148CBA",
  "M. aurum/HC" = "#98E0B1",
  "M. aurum/IS" = "#21A951"
)

alpha_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11),
    legend.key.size = unit(0.6, "cm"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 10, angle = 25, hjust = 1),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5)
  )

sample_type_faith <- ggplot(alpha_div_metadata, aes(x = experimental_group, y = faith_pd, fill = experimental_group)) +
  geom_boxplot(width = 0.7, outlier.shape = NA, linewidth = 0.7) +
  geom_jitter(width = 0.12, size = 2.2, alpha = 0.9) +
  scale_fill_manual(
  values = group_colors,
  name = "Group",
  labels = c(
    "BBS/HC",
    "BBS/IS",
    expression(italic("M. aurum") ~ "/HC"),
    expression(italic("M. aurum") ~ "/IS")
  )
) +
  scale_x_discrete(
    labels = c(
      "BBS/HC",
      "BBS/IS",
      expression(italic("M. aurum") ~ "/HC"),
      expression(italic("M. aurum") ~ "/IS")
    )
  ) +
  labs(x = NULL, y = "Faith's diversity index") +
  alpha_theme

sample_type_evenness <- ggplot(alpha_div_metadata, aes(x = experimental_group, y = pielou_evenness, fill = experimental_group)) +
  geom_boxplot(width = 0.7, outlier.shape = NA, linewidth = 0.7) +
  geom_jitter(width = 0.12, size = 2.2, alpha = 0.9) +
  scale_fill_manual(
  values = group_colors,
  name = "Group",
  labels = c(
    "BBS/HC",
    "BBS/IS",
    expression(italic("M. aurum") ~ "/HC"),
    expression(italic("M. aurum") ~ "/IS")
  )
) +
  scale_x_discrete(
    labels = c(
      "BBS/HC",
      "BBS/IS",
      expression(italic("M. aurum") ~ "/HC"),
      expression(italic("M. aurum") ~ "/IS")
    )
  ) +
  labs(x = NULL, y = "Pielou's evenness") +
  alpha_theme

sample_type_shannon <- ggplot(alpha_div_metadata, aes(x = experimental_group, y = shannon_entropy, fill = experimental_group)) +
  geom_boxplot(width = 0.7, outlier.shape = NA, linewidth = 0.7) +
  geom_jitter(width = 0.12, size = 2.2, alpha = 0.9) +
  scale_fill_manual(
  values = group_colors,
  name = "Group",
  labels = c(
    "BBS/HC",
    "BBS/IS",
    expression(italic("M. aurum") ~ "/HC"),
    expression(italic("M. aurum") ~ "/IS")
  )
) +
  scale_x_discrete(
    labels = c(
      "BBS/HC",
      "BBS/IS",
      expression(italic("M. aurum") ~ "/HC"),
      expression(italic("M. aurum") ~ "/IS")
    )
  ) +
  labs(x = NULL, y = "Shannon Diversity") +
  alpha_theme

sample_type_obs_feat <- ggplot(alpha_div_metadata, aes(x = experimental_group, y = observed_features, fill = experimental_group)) +
  geom_boxplot(width = 0.7, outlier.shape = NA, linewidth = 0.7) +
  geom_jitter(width = 0.12, size = 2.2, alpha = 0.9) +
  scale_fill_manual(
  values = group_colors,
  name = "Group",
  labels = c(
    "BBS/HC",
    "BBS/IS",
    expression(italic("M. aurum") ~ "/HC"),
    expression(italic("M. aurum") ~ "/IS")
  )
) +
  scale_x_discrete(
    labels = c(
      "BBS/HC",
      "BBS/IS",
      expression(italic("M. aurum") ~ "/HC"),
      expression(italic("M. aurum") ~ "/IS")
    )
  ) +
  labs(x = NULL, y = "Observed ASV") +
  alpha_theme

alpha_div_sample_type <- ggarrange(
  sample_type_faith, sample_type_evenness,
  sample_type_shannon, sample_type_obs_feat,
  ncol = 2, nrow = 2,
  labels = c("A", "B", "C", "D"),
  common.legend = TRUE,
  legend = "bottom"
)

alpha_div_sample_type

ggsave(
  filename = "R_figures/alpha_div_sample_type.svg",
  plot = alpha_div_sample_type,
  device = svglite::svglite,
  width = 8,
  height = 8,
  units = "in"
)

# Treatment did not see any significance, but Dr. Lowry is very interesting in the interactions of HC and IS on 082725. Continue analyzing these comparisons for the generating the figures.
```
## Making figure- Condition since we see something interesting with Condition from the Treatment 
```{r}
# -----------------------------
# 1. Clean Condition factor
# -----------------------------
alpha_div_metadata$Condition <- factor(
  as.character(alpha_div_metadata$Condition),
  levels = c("HC", "IS")
)

condition_colors <- c(
  "HC" = "#B3E1F0",
  "IS" = "#148CBA"
)

alpha_theme <- theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11),
    legend.key.size = unit(0.6, "cm"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10)
  )

# -----------------------------
# 2. Build plots
# -----------------------------
condition_faith <- ggplot(alpha_div_metadata, aes(x = Condition, y = faith_pd, fill = Condition)) +
  geom_boxplot(width = 0.7, outlier.shape = NA, linewidth = 0.7) +
  geom_jitter(width = 0.12, size = 2.2, alpha = 0.9) +
  scale_fill_manual(values = condition_colors, name = "Condition") +
  labs(x = NULL, y = "Faith's diversity index") +
  alpha_theme

condition_evenness <- ggplot(alpha_div_metadata, aes(x = Condition, y = pielou_evenness, fill = Condition)) +
  geom_boxplot(width = 0.7, outlier.shape = NA, linewidth = 0.7) +
  geom_jitter(width = 0.12, size = 2.2, alpha = 0.9) +
  scale_fill_manual(values = condition_colors, name = "Condition") +
  labs(x = NULL, y = "Pielou's evenness") +
  alpha_theme

condition_shannon <- ggplot(alpha_div_metadata, aes(x = Condition, y = shannon_entropy, fill = Condition)) +
  geom_boxplot(width = 0.7, outlier.shape = NA, linewidth = 0.7) +
  geom_jitter(width = 0.12, size = 2.2, alpha = 0.9) +
  scale_fill_manual(values = condition_colors, name = "Condition") +
  labs(x = NULL, y = "Shannon diversity") +
  alpha_theme

condition_obs_feat <- ggplot(alpha_div_metadata, aes(x = Condition, y = observed_features, fill = Condition)) +
  geom_boxplot(width = 0.7, outlier.shape = NA, linewidth = 0.7) +
  geom_jitter(width = 0.12, size = 2.2, alpha = 0.9) +
  scale_fill_manual(values = condition_colors, name = "Condition") +
  labs(x = NULL, y = "Observed ASV") +
  alpha_theme

# -----------------------------
# 3. Combine plots
# -----------------------------
alpha_div_condition <- ggarrange(
  condition_faith, condition_evenness,
  condition_shannon, condition_obs_feat,
  ncol = 2, nrow = 2,
  labels = c("A", "B", "C", "D"),
  common.legend = TRUE,
  legend = "bottom"
)

alpha_div_condition

ggsave(
  filename = "R_figures/alpha_div_condition.svg",
  plot = alpha_div_condition,
  device = svglite::svglite,
  width = 8,
  height = 8,
  units = "in"
)
```
################################################################################
# Beta diversity- ADONIS and PERMANOVAS 
################################################################################
## Import beta diversity data from QIIME2 output
```{r}
# note - you can import the ordination/PCoA results from QIIME2 instead of the distance matrices, which require recalculation of ordinations
# but, extracting the graphable data from the PCoA results is a little more complex, and importing the distance matrices/recalculating the ordination is very transparent and also teaches how to create ordinations in R

bray_curtis_dm <- read_tsv("output_for_R/bray_curtis_distance_matrix.tsv")
jaccard_dm <- read_tsv("output_for_R/jaccard_distance_matrix.tsv")
uw_unifrac_dm <- read_tsv("output_for_R/unweighted_unifrac_distance_matrix.tsv")
w_unifrac_dm <- read_tsv("output_for_R/weighted_unifrac_distance_matrix.tsv")

# notice that the sample names are in the first column. We want them to be the rownames
bray_curtis_dm <- column_to_rownames(bray_curtis_dm, var = "...1")
jaccard_dm <- column_to_rownames(jaccard_dm, var = "...1")
uw_unifrac_dm <- column_to_rownames(uw_unifrac_dm, var = "...1")
w_unifrac_dm <- column_to_rownames(w_unifrac_dm, var = "...1")

# import metadata
metadata <- read_tsv("Maurum_only_metadata.tsv")
```
## PCoA plots and PERMANOVA for Treatment
```{r}
# create ordinations from our distance matrices
bray_curtis_pcoa <- cmdscale(bray_curtis_dm, eig = TRUE)
jaccard_pcoa <- cmdscale(jaccard_dm, eig = TRUE)
uw_unifrac_pcoa <- cmdscale(uw_unifrac_dm, eig = TRUE)
w_unifrac_pcoa <- cmdscale(w_unifrac_dm, eig = TRUE)

# create a graphable data frame from our pcoa results 
bray_curtis_df <- data.frame(bray_curtis_pcoa$points)
jaccard_df <- data.frame(jaccard_pcoa$points)
uw_unifrac_df <- data.frame(uw_unifrac_pcoa$points)
w_unifrac_df <- data.frame(w_unifrac_pcoa$points)

# extract the % variation for PCoA axes 1 and 2
bray_curtis_eg <- bray_curtis_pcoa$eig
bray_curtis_var <- bray_curtis_eg / sum(bray_curtis_eg[bray_curtis_eg > 0]) * 100

jaccard_eg <- jaccard_pcoa$eig
jaccard_var <- jaccard_eg / sum(jaccard_eg[jaccard_eg > 0]) * 100

uw_unifrac_eg <- uw_unifrac_pcoa$eig
uw_unifrac_var <- uw_unifrac_eg / sum(uw_unifrac_eg[uw_unifrac_eg > 0]) * 100

w_unifrac_eg <- w_unifrac_pcoa$eig
w_unifrac_var <- w_unifrac_eg / sum(w_unifrac_eg[w_unifrac_eg > 0]) * 100

# rename the columns in our data frames
colnames(bray_curtis_df) <- c("PCoA1", "PCoA2")
colnames(jaccard_df) <- c("PCoA1", "PCoA2")
colnames(uw_unifrac_df) <- c("PCoA1", "PCoA2")
colnames(w_unifrac_df) <- c("PCoA1", "PCoA2")

# add metadata to our data frames
bray_curtis_df$sample_name <- rownames(bray_curtis_df)
bray_curtis_df <- bray_curtis_df %>% 
  left_join(metadata, bray_curtis_df, by = c("sample_name" = "SampleID"))

jaccard_df$sample_name <- rownames(jaccard_df)
jaccard_df <- jaccard_df %>% 
  left_join(metadata, jaccard_df, by = c("sample_name" = "SampleID"))

uw_unifrac_df$sample_name <- rownames(uw_unifrac_df)
uw_unifrac_df <- uw_unifrac_df %>% 
  left_join(metadata, uw_unifrac_df, by = c("sample_name" = "SampleID"))

w_unifrac_df$sample_name <- rownames(w_unifrac_df)
w_unifrac_df <- w_unifrac_df %>% 
  left_join(metadata, w_unifrac_df, by = c("sample_name" = "SampleID"))

# Ensure the same sample order in metadata as in distance matrices before calculating stats
bray_curtis_df <- column_to_rownames(bray_curtis_df, var = "sample_name")
bray_curtis_df <- bray_curtis_df[match(rownames(as.matrix(bray_curtis_df)), rownames(bray_curtis_df)), ]
all(rownames(bray_curtis_df) == rownames(as.matrix(bray_curtis_dm)))  # checks previous line, should be TRUE

jaccard_df <- column_to_rownames(jaccard_df, var = "sample_name")
jaccard_df <- jaccard_df[match(rownames(as.matrix(jaccard_df)), rownames(jaccard_df)), ]
all(rownames(jaccard_df) == rownames(as.matrix(jaccard_dm)))  # checks previous line, should be TRUE

uw_unifrac_df <- column_to_rownames(uw_unifrac_df, var = "sample_name")
uw_unifrac_df <- uw_unifrac_df[match(rownames(as.matrix(uw_unifrac_df)), rownames(uw_unifrac_df)), ]
all(rownames(uw_unifrac_df) == rownames(as.matrix(uw_unifrac_dm)))  # checks previous line, should be TRUE

w_unifrac_df <- column_to_rownames(w_unifrac_df, var = "sample_name")
w_unifrac_df <- w_unifrac_df[match(rownames(as.matrix(w_unifrac_df)), rownames(w_unifrac_df)), ]
all(rownames(w_unifrac_df) == rownames(as.matrix(w_unifrac_dm)))  # checks previous line, should be TRUE

# calculate PERMANOVA statistics between sample type- Vegan Adonis
bray_curtis_sample_type_perm <- adonis2(bray_curtis_dm ~ experimental_group, data = bray_curtis_df)
bray_curtis_sample_type_r2 <- round(bray_curtis_sample_type_perm$R2[1], 3)
bray_curtis_sample_type_pval <- signif(bray_curtis_sample_type_perm$`Pr(>F)`[1], 3)

jaccard_sample_type_perm <- adonis2(jaccard_dm ~ experimental_group, data = jaccard_df)
jaccard_sample_type_r2 <- round(jaccard_sample_type_perm$R2[1], 3)
jaccard_sample_type_pval <- signif(jaccard_sample_type_perm$`Pr(>F)`[1], 3)

uw_unifrac_sample_type_perm <- adonis2(uw_unifrac_dm ~ experimental_group, data = uw_unifrac_df)
uw_unifrac_sample_type_r2 <- round(uw_unifrac_sample_type_perm$R2[1], 3)
uw_unifrac_sample_type_pval <- signif(uw_unifrac_sample_type_perm$`Pr(>F)`[1], 3)

w_unifrac_sample_type_perm <- adonis2(w_unifrac_dm ~ experimental_group, data = w_unifrac_df)
w_unifrac_sample_type_r2 <- round(w_unifrac_sample_type_perm$R2[1], 3)
w_unifrac_sample_type_pval <- signif(w_unifrac_sample_type_perm$`Pr(>F)`[1], 3)

bray_curtis_df$experimental_group <- dplyr::recode(
  bray_curtis_df$experimental_group,
  "bbs.HC"    = "BBS/HC",
  "bbs.IS"    = "BBS/IS",
  "maurum.HC" = "M. aurum/HC",
  "maurum.IS" = "M. aurum/IS"
)

jaccard_df$experimental_group <- dplyr::recode(
  jaccard_df$experimental_group,
  "bbs.HC"    = "BBS/HC",
  "bbs.IS"    = "BBS/IS",
  "maurum.HC" = "M. aurum/HC",
  "maurum.IS" = "M. aurum/IS"
)

uw_unifrac_df$experimental_group <- dplyr::recode(
  uw_unifrac_df$experimental_group,
  "bbs.HC"    = "BBS/HC",
  "bbs.IS"    = "BBS/IS",
  "maurum.HC" = "M. aurum/HC",
  "maurum.IS" = "M. aurum/IS"
)

w_unifrac_df$experimental_group <- dplyr::recode(
  w_unifrac_df$experimental_group,
  "bbs.HC"    = "BBS/HC",
  "bbs.IS"    = "BBS/IS",
  "maurum.HC" = "M. aurum/HC",
  "maurum.IS" = "M. aurum/IS"
)

# graph sample type PCoA plots
sample_type_bray_curtis <- ggplot(bray_curtis_df, aes(x = PCoA1, y = PCoA2, color = experimental_group)) +
  geom_point() +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black")
  )+
  stat_ellipse(type = "t", level = 0.95) +
  xlab(paste0("PCoA1 (", round(bray_curtis_var[1], 1), "%)")) +
  ylab(paste0("PCoA2 (", round(bray_curtis_var[2], 1), "%)")) +
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("PERMANOVA: R² = ", bray_curtis_sample_type_r2, 
                          ", p = ", bray_curtis_sample_type_pval),
           hjust = 1.1, vjust = -0.5,
           size = 4) +
  scale_color_manual(
  values = c(
    "BBS/HC"        = "#B3E1F0",
    "BBS/IS"        = "#148CBA",
    "M. aurum/HC"   = "#98E0B1",
    "M. aurum/IS"   = "#21A951"
  ),
labels = c(
    "BBS/HC",
    "BBS/IS",
    expression(italic("M. aurum") ~ "/HC"),
    expression(italic("M. aurum") ~ "/IS")
  )
) +
theme(
  legend.title = element_text(size = 13, face = "bold"),
  legend.text  = element_text(size = 12),
  legend.key.size = unit(0.7, "cm")
) +
  labs(color = "Experimental Group", title = "Bray-Curtis")
sample_type_bray_curtis

sample_type_jaccard <- ggplot(jaccard_df, aes(x = PCoA1, y = PCoA2, color = experimental_group)) +
  geom_point() +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black")
  )+
  stat_ellipse(type = "t", level = 0.95) +
  xlab(paste0("PCoA1 (", round(jaccard_var[1], 1), "%)")) +
  ylab(paste0("PCoA2 (", round(jaccard_var[2], 1), "%)")) +
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("PERMANOVA: R² = ", jaccard_sample_type_r2, 
                          ", p = ", jaccard_sample_type_pval),
           hjust = 1.1, vjust = -0.5,
           size = 4) +
  scale_color_manual(
  values = c(
    "BBS/HC"        = "#B3E1F0",
    "BBS/IS"        = "#148CBA",
    "M. aurum/HC"   = "#98E0B1",
    "M. aurum/IS"   = "#21A951"
  ),
labels = c(
    "BBS/HC",
    "BBS/IS",
    expression(italic("M. aurum") ~ "/HC"),
    expression(italic("M. aurum") ~ "/IS")
  )
) +
theme(
  legend.title = element_text(size = 13, face = "bold"),
  legend.text  = element_text(size = 12),
  legend.key.size = unit(0.7, "cm")
) +
  labs(fill = "Experimental Group", title = "Jaccard")
sample_type_jaccard

sample_type_uw_unifrac <- ggplot(uw_unifrac_df, aes(x = PCoA1, y = PCoA2, color = experimental_group)) +
  geom_point() +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black")
  )+
  stat_ellipse(type = "t", level = 0.95) +
  xlab(paste0("PCoA1 (", round(uw_unifrac_var[1], 1), "%)")) +
  ylab(paste0("PCoA2 (", round(uw_unifrac_var[2], 1), "%)")) +
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("PERMANOVA: R² = ", uw_unifrac_sample_type_r2, 
                          ", p = ", uw_unifrac_sample_type_pval),
           hjust = 1.1, vjust = -0.5,
           size = 4) +
  scale_color_manual(
  values = c(
    "BBS/HC"        = "#B3E1F0",
    "BBS/IS"        = "#148CBA",
    "M. aurum/HC"   = "#98E0B1",
    "M. aurum/IS"   = "#21A951"
  ),
labels = c(
    "BBS/HC",
    "BBS/IS",
    expression(italic("M. aurum") ~ "/HC"),
    expression(italic("M. aurum") ~ "/IS")
  )
) +
theme(
  legend.title = element_text(size = 13, face = "bold"),
  legend.text  = element_text(size = 12),
  legend.key.size = unit(0.7, "cm")
) +
  labs(fill = "Experimental Group", title = "Unweighted Unifrac")
sample_type_uw_unifrac

sample_type_w_unifrac <- ggplot(w_unifrac_df, aes(x = PCoA1, y = PCoA2, color = experimental_group)) +
  geom_point() +
  theme_bw() +
theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black")
  )+
  stat_ellipse(type = "t", level = 0.95) +
  xlab(paste0("PCoA1 (", round(w_unifrac_var[1], 1), "%)")) +
  ylab(paste0("PCoA2 (", round(w_unifrac_var[2], 1), "%)")) +
  annotate("text", x = Inf, y = -Inf, 
           label = paste0("PERMANOVA: R² = ", w_unifrac_sample_type_r2, 
                          ", p = ", w_unifrac_sample_type_pval),
           hjust = 1.1, vjust = -0.5,
           size = 4) +
 scale_color_manual(
  values = c(
    "BBS/HC"        = "#B3E1F0",
    "BBS/IS"        = "#148CBA",
    "M. aurum/HC"   = "#98E0B1",
    "M. aurum/IS"   = "#21A951"
  ),
 labels = c(
    "BBS/HC",
    "BBS/IS",
    expression(italic("M. aurum") ~ "/HC"),
    expression(italic("M. aurum") ~ "/IS")
  )
) +
theme(
  legend.title = element_text(size = 13, face = "bold"),
  legend.text  = element_text(size = 12),
  legend.key.size = unit(0.7, "cm")
) +
  labs(fill = "Sample Type", title = "Weighted Unifrac")
sample_type_w_unifrac

# merge plots and save
beta_div_sample_type <- ggarrange(sample_type_bray_curtis, sample_type_jaccard,
                                  sample_type_uw_unifrac, sample_type_w_unifrac,
                                  ncol = 2, nrow = 2,
                                  labels = c("A","B","C","D"),
                                  common.legend = TRUE)
beta_div_sample_type

ggsave(
  filename = "R_figures/beta_div_sample_type.svg",
  plot = beta_div_sample_type,
  device = svglite::svglite,
  width = 8,
  height = 8,
  units = "in"
)
```
## PCoA plots and PERMANOVA for Condition
``` {r}
# make sure Condition is ordered
bray_curtis_df$Condition <- factor(bray_curtis_df$Condition, levels = c("HC", "IS"))
jaccard_df$Condition <- factor(jaccard_df$Condition, levels = c("HC", "IS"))
uw_unifrac_df$Condition <- factor(uw_unifrac_df$Condition, levels = c("HC", "IS"))
w_unifrac_df$Condition <- factor(w_unifrac_df$Condition, levels = c("HC", "IS"))

# -----------------------------
# 1. PERMANOVA using full data
# -----------------------------
bray_curtis_condition_perm <- adonis2(bray_curtis_dm ~ Condition, data = bray_curtis_df)
bray_curtis_condition_r2 <- round(bray_curtis_condition_perm$R2[1], 2)
bray_curtis_condition_pval <- signif(bray_curtis_condition_perm$`Pr(>F)`[1], 2)

jaccard_condition_perm <- adonis2(jaccard_dm ~ Condition, data = jaccard_df)
jaccard_condition_r2 <- round(jaccard_condition_perm$R2[1], 2)
jaccard_condition_pval <- signif(jaccard_condition_perm$`Pr(>F)`[1], 2)

uw_unifrac_condition_perm <- adonis2(uw_unifrac_dm ~ Condition, data = uw_unifrac_df)
uw_unifrac_condition_r2 <- round(uw_unifrac_condition_perm$R2[1], 2)
uw_unifrac_condition_pval <- signif(uw_unifrac_condition_perm$`Pr(>F)`[1], 2)

w_unifrac_condition_perm <- adonis2(w_unifrac_dm ~ Condition, data = w_unifrac_df)
w_unifrac_condition_r2 <- round(w_unifrac_condition_perm$R2[1], 2)
w_unifrac_condition_pval <- signif(w_unifrac_condition_perm$`Pr(>F)`[1], 2)

# -----------------------------
# 2. Shared color scale
# -----------------------------
condition_colors <- scale_color_manual(
  values = c(
    "HC" = "#B3E1F0",
    "IS" = "#148CBA"
  ),
  labels = c(
    "HC" = "HC",
    "IS" = "IS"
  )
)

# -----------------------------
# 3. Plot function
# -----------------------------
make_condition_plot <- function(df, var, r2, pval, title_text) {
  ggplot(df, aes(x = PCoA1, y = PCoA2, color = Condition)) +
    geom_point(size = 2.5) +
    theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black")
  )+
    stat_ellipse(type = "t", level = 0.95, linewidth = 0.8) +
    xlab(paste0("PCoA1 (", round(var[1], 1), "%)")) +
    ylab(paste0("PCoA2 (", round(var[2], 1), "%)")) +
    annotate(
      "text",
      x = Inf, y = -Inf,
      label = paste0("PERMANOVA: R² = ", r2, ", p = ", pval),
      hjust = 1.1, vjust = -0.5,
      size = 4
    ) +
theme(
  legend.title = element_text(size = 13, face = "bold"),
  legend.text  = element_text(size = 12),
  legend.key.size = unit(0.7, "cm")
) +
    labs(color = "Condition", title = title_text) +
    condition_colors
}

# -----------------------------
# 4. Build the four plots
# -----------------------------
condition_bray_curtis <- make_condition_plot(
  bray_curtis_df, bray_curtis_var,
  bray_curtis_condition_r2, bray_curtis_condition_pval,
  "Bray-Curtis"
)

condition_jaccard <- make_condition_plot(
  jaccard_df, jaccard_var,
  jaccard_condition_r2, jaccard_condition_pval,
  "Jaccard"
)

condition_uw_unifrac <- make_condition_plot(
  uw_unifrac_df, uw_unifrac_var,
  uw_unifrac_condition_r2, uw_unifrac_condition_pval,
  "Unweighted UniFrac"
)

condition_w_unifrac <- make_condition_plot(
  w_unifrac_df, w_unifrac_var,
  w_unifrac_condition_r2, w_unifrac_condition_pval,
  "Weighted UniFrac"
)

# -----------------------------
# 5. Combine
# -----------------------------
beta_div_condition <- ggarrange(
  condition_bray_curtis, condition_jaccard,
  condition_uw_unifrac, condition_w_unifrac,
  ncol = 2, nrow = 2,
  labels = c("A", "B", "C", "D"),
  common.legend = TRUE,
  legend = "top"
)

ggsave(
  filename = "R_figures/beta_div_condition.svg",
  plot = beta_div_condition,
  device = svglite::svglite,
  width = 8,
  height = 8,
  units = "in"
)
```
## Statistical Analysis: PERMDISP
```{r}
bray_curtis_dist <- as.dist(bray_curtis_dm)

stopifnot(attr(bray_curtis_dist, "Size") == nrow(bray_curtis_df))

bd_bray <- betadisper(bray_curtis_dist, bray_curtis_df$experimental_group)
perm_bray <- permutest(bd_bray, permutations = 99)

perm_bray
# p> 0.05 (0.98), dispersion is similar across groups, PERMANOVA reflects true community differences

jaccard_dist <- as.dist(jaccard_dm)

bd_jaccard <- betadisper(jaccard_dist, jaccard_df$experimental_group)
perm_jaccard <- permutest(bd_jaccard, permutations = 99)

perm_jaccard
# p= 0.47
uw_unifrac_dist <- as.dist(uw_unifrac_dm)

bd_uw <- betadisper(uw_unifrac_dist, uw_unifrac_df$experimental_group)
perm_uw <- permutest(bd_uw, permutations = 99)

perm_uw
# p= 0.23
w_unifrac_dist <- as.dist(w_unifrac_dm)

bd_w <- betadisper(w_unifrac_dist, w_unifrac_df$experimental_group)
perm_w <- permutest(bd_w, permutations = 99)

perm_w
# p= 0.99
```
### Export non-rarefied table ####
```{r}
table <- read_qza("data/maurum_table.qza")$data

# get names
taxonomy <- read_qza("data/silva-taxonomy.qza")$data
asvs <- rownames(table)
taxa <- vector(length=length(asvs))

i <- 1
for (a in asvs){
    taxa[i] <- taxonomy[taxonomy$Feature.ID==a,"Taxon"]
    i <- i + 1
}

taxa_clean <- give_nickname(taxa, one_name = TRUE)
taxa_phylum <- give_nickname(taxa, one_name = TRUE, phylum = TRUE)
#rarefied_table_tax <- rarefied_table
#rownames(rarefied_table_tax) <- taxa_clean
#rownames(metadata) <- metadata$SampleID

t.table <- t(table)
rs <- rowSums(t.table)
df <- as.data.frame(rs)
df <- df %>%
  dplyr::rename( "otus" = "rs" ) #had to call dplyer package for this

write.csv(df, "data/hf_otus.csv", row.names=TRUE)
```
################################################################################
# Taxonomy
################################################################################
## Import feature table and taxonomy data from QIIME2 output
```{r}
tabulated_results <- read_tsv("output_for_R/tabulated_results.tsv")
```
## Format, convert taxonomy to relative abundance, and group at taxonomic levels
```{r}
# format data for making a taxa plot
# get rid of #q2:types row
tabulated_results <- tabulated_results[2:nrow(tabulated_results),]

# remove non-numeric columns
# we will filter out Taxon for now so we can calculate relative abundances, but will add it back later
tabulated_results_filt <- tabulated_results %>% 
  select(-all_of(c("Sequence","Confidence","Taxon")))
# we keep id and make it as the row name
tabulated_results_filt <- column_to_rownames(tabulated_results_filt, var = "id")

# calculate relative abundances
tabulated_results_filt[] <- lapply(tabulated_results_filt, function(x) as.numeric(as.character(x)))
rel_abundance <- apply(tabulated_results_filt, 2, function(x) x / sum(x))

# check that each column adds up to 1
colSums(rel_abundance)

# re-add taxonomy to relative abundance table
rel_abundance <- data.frame(rel_abundance, check.names = FALSE)
rel_abundance$id <- rownames(rel_abundance)

tabulated_results_merge <- tabulated_results %>% # create a tabulated results for merging
  select(all_of(c("id","Taxon")))

rel_abundance_taxa <- merge(rel_abundance, tabulated_results_merge, by = "id", all.x = TRUE)

# we now have a table with the per-sample relative abundances of each ASV (along with their taxonomic assignments)

# Taxon column data is messy - let's split the data up into taxonomic ranks
rel_abundance_taxa_split <- rel_abundance_taxa %>% 
  separate(Taxon, 
           into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"),
           sep = ";\\s*",  # splits on semicolon and optional space
           fill = "right", # fills missing columns with NA if some rows are shorter
           remove = TRUE) # remove original "Taxon" column too

# some ranks have a value even though they are unidentified (e.g., g__, s__)
# let's change those to NAs
rel_abundance_taxa_split[rel_abundance_taxa_split == "d__"] <- NA
rel_abundance_taxa_split[rel_abundance_taxa_split == "p__"] <- NA
rel_abundance_taxa_split[rel_abundance_taxa_split == "c__"] <- NA
rel_abundance_taxa_split[rel_abundance_taxa_split == "o__"] <- NA
rel_abundance_taxa_split[rel_abundance_taxa_split == "f__"] <- NA
rel_abundance_taxa_split[rel_abundance_taxa_split == "g__"] <- NA
rel_abundance_taxa_split[rel_abundance_taxa_split == "s__"] <- NA

# create a column with the lowest identified taxonomic level for all ASVs
rel_abundance_taxa_lowest <- rel_abundance_taxa_split %>% 
  rowwise() %>%
  mutate(LowestTaxLevel = {
    tax_levels <- c(Domain, Phylum, Class, Order, Family, Genus, Species)
    last_non_na <- tail(na.omit(tax_levels), 1)
    ifelse(length(last_non_na) > 0, last_non_na, NA)
  }) %>%
  ungroup()


# group by the lowest taxonomic level
rel_abundance_taxa_lowest <- rel_abundance_taxa_lowest %>%
dplyr::mutate(LowestTaxLevel = ifelse(is.na(LowestTaxLevel), "Unassigned", LowestTaxLevel)) %>% # convert NAs to "Unassigned"
dplyr::group_by(LowestTaxLevel) %>%
dplyr::summarise(
dplyr::across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  )
# save the file
saveRDS(rel_abundance_taxa_lowest, "R_output_data/rel_abundance_taxa_lowest.RDS")

# group at level 5/Family level
rel_abundance_taxa_family <- rel_abundance_taxa_split %>%
  dplyr::mutate(Family = ifelse(is.na(Family), "Unassigned", Family)) %>%
  dplyr::group_by(Family) %>%
  dplyr::summarise(
    dplyr::across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  )

# save the file
saveRDS(rel_abundance_taxa_family, "R_output_data/rel_abundance_taxa_family.RDS")

# group at level 6/Genus level 
rel_abundance_taxa_genus <- rel_abundance_taxa_split %>%
  dplyr::mutate(Genus = ifelse(is.na(Genus), "Unassigned", Genus)) %>%
  dplyr::group_by(Genus) %>%
  dplyr::summarise(
    dplyr::across(where(is.numeric), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  )

# save the file
saveRDS(rel_abundance_taxa_genus, "R_output_data/rel_abundance_taxa_genus.RDS")

# this allows you to have either the lowest-identified level or collapsed taxonomic tables 
```
## Generating Taxaplot
```{r}
# Build your phyloseq
pseq <- qiime2R::qza_to_phyloseq(
  feature  = "maurum_table.qza",
  taxonomy = "silva-taxonomy.qza",
  metadata = "Maurum_only_metadata.tsv"
)

taxonomy.table <- as.data.frame(tax_table(pseq))

phylum_name_updates <- c(
  "Firmicutes"      = "Bacillota",
  "Bacteroidetes"   = "Bacteroidota",
  "Proteobacteria"  = "Pseudomonadota",
  "Actinobacteria"  = "Actinomycetota",
  "Fusobacteria"    = "Fusobacteriota",
  "Lentisphaerae"   = "Lentisphaerota",
  "Tenericutes"     = "Mycoplasmatota",
  "Synergistetes"   = "Synergistota",
  "Verrucomicrobia" = "Verrucomicrobiota"
)

make_microshades_phylum <- function(pseq, subgroup) {

  # Rename metadata column "Sample" if needed
  sd <- as.data.frame(phyloseq::sample_data(pseq))
  if ("Sample" %in% names(sd)) {
    names(sd)[names(sd) == "Sample"] <- "Sample_meta"
    phyloseq::sample_data(pseq) <- phyloseq::sample_data(sd)
  }

  # Prep microshades dataframe
  mdf_prep <- microshades::prep_mdf(pseq, subgroup_level = subgroup)

  # Update phylum names for display
  if ("Phylum" %in% names(mdf_prep)) {
    mdf_prep$Phylum <- dplyr::recode(
      as.character(mdf_prep$Phylum),
      !!!phylum_name_updates,
      .default = as.character(mdf_prep$Phylum)
    )
  }

bad_taxa <- "uncultured|unknown|unclassified|ambiguous|metagenome"

# Remove unwanted family labels BEFORE microshades makes the key
mdf_prep <- mdf_prep %>%
  dplyr::filter(
    !is.na(Phylum),
    Phylum != "",
    !grepl(bad_taxa, Phylum, ignore.case = TRUE),
    !is.na(.data[[subgroup]]),
    .data[[subgroup]] != "",
    !grepl(bad_taxa, .data[[subgroup]], ignore.case = TRUE)
  )

# Order phyla by total abundance, greatest to least
selected_phyla <- mdf_prep %>%
  dplyr::group_by(Phylum) %>%
  dplyr::summarise(
    total_abundance = sum(Abundance, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(dplyr::desc(total_abundance)) %>%
  dplyr::slice_head(n = 5) %>%
  dplyr::pull(Phylum)

  # Build microshades color/data frames
  co <- microshades::create_color_dfs(
    mdf_prep,
    selected_groups = selected_phyla,
    group_level     = "Phylum",
    subgroup_level  = subgroup,
    top_n_subgroups = 4,
    cvd             = TRUE
  )

  mdf_GP <- co$mdf
  cdf_GP <- co$cdf

  # Order ONLY the key/legend phylum groups, not the plotted bars
  final_group_order <- c(
    selected_phyla,
    setdiff(unique(as.character(cdf_GP$group)), selected_phyla)
  )

  cdf_GP$group <- factor(
    as.character(cdf_GP$group),
    levels = final_group_order
  )

  cdf_GP <- cdf_GP %>%
    dplyr::arrange(group)

  # Custom legend/key
  GP_legend <- custom_legend(
    mdf_GP, cdf_GP,
    legend_key_size  = unit(0.4, "cm"),
    legend_text_size = 10,
    subgroup_level   = subgroup
  )

  # Rename facet labels
  mdf_GP$experimental_group <- dplyr::recode(
    mdf_GP$experimental_group,
    "bbs.HC"    = "BBS/HC",
    "bbs.IS"    = "BBS/IS",
    "maurum.HC" = "M. aurum/HC",
    "maurum.IS" = "M. aurum/IS"
  )

  # Plot
  p <- microshades::plot_microshades(
    mdf_GP, cdf_GP,
    x = "Sample",
    y = "Abundance",
    group_label = paste("Phylum", subgroup)
  ) +
    scale_y_continuous(
      labels = scales::percent,
      expand = expansion(0)
    ) +
    labs(y = "Relative abundance") +
    theme(legend.position = "none") +
    facet_grid(
      ~ experimental_group,
      scales = "free_x",
      space = "free_x",
      labeller = as_labeller(c(
        "BBS/HC"      = "BBS/HC",
        "BBS/IS"      = "BBS/IS",
        "M. aurum/HC" = "italic('M. aurum')*'/HC'",
        "M. aurum/IS" = "italic('M. aurum')*'/IS'"
      ), label_parsed)
    ) +
    theme(
      axis.title.y = element_text(size = 14, face = "bold"),
      axis.title.x = element_text(size = 14, face = "bold"),
      axis.text.y  = element_text(size = 11),
      axis.text.x  = element_text(size = 11),
      strip.text   = element_text(size = 13, face = "bold")
    )

  cowplot::plot_grid(p, GP_legend, rel_widths = c(1, 0.30))
}

# Run
multi <- make_microshades_phylum(pseq, "Family")

# Save
ggsave(
  filename = "R_figures/taxa_barplot_experimentalgroup.svg",
  plot = multi,
  device = svglite::svglite,
  width = 20,
  height = 8,
  units = "in"
)
```
################################################################################
# Differential Abundance- ASV, Relative Abundance, and ANCOMBC 
################################################################################
## ASV Level 
```{r}
# Parse taxonomy into ranks- Order, Family, Genus, Species
metadata <- read_q2metadata("Maurum_only_metadata.tsv")
metadata <- metadata[metadata$SampleID != "#q2:types", ]
rownames(metadata) <- metadata$SampleID

asv_table <- read_qza("maurum_table.qza")$data
taxonomy_parsed <- read_qza("silva-taxonomy.qza")$data %>%
  parse_taxonomy()

taxonomy_parsed <- taxonomy_parsed[match(rownames(asv_table), rownames(taxonomy_parsed)), ]

ps_asv <- phyloseq(
  otu_table(as.matrix(asv_table), taxa_are_rows = TRUE),
  tax_table(as.matrix(taxonomy_parsed)),
  sample_data(metadata)
)

rank_names(ps_asv)
#Interpretation of taxonomy coverage is great ~99% for Order and Family. ~82% for Genus. ~50% for species. Need to be careful at interpret it since many ASVs will be NA, uncultured, or ambiguous
```
## Relative abundance Treatment: Phylum
```{r}
#Make sure to do parsed taxonomy ranks from the ASV code above before doing this step

#naming a critical object
pseq <- ps_asv
# ---------------------------------
# 1. Build clean sample group labels
# ---------------------------------
sample_data(pseq)$experimental_group <- factor(
  as.character(sample_data(pseq)$experimental_group),
  levels = c("bbs.HC", "bbs.IS", "maurum.HC", "maurum.IS"),
  labels = c("BBS/HC", "BBS/IS", "M. aurum/HC", "M. aurum/IS")
)

# ---------------------------------
# 2. Agglomerate at phylum level
# ---------------------------------
pseq_phylum <- tax_glom(pseq, taxrank = "Phylum", NArm = FALSE)

# ---------------------------------
# 3. Transform to relative abundance
# ---------------------------------
pseq_phylum_rel <- transform_sample_counts(pseq_phylum, function(x) x / sum(x))

# ---------------------------------
# 4. Melt to long format
# ---------------------------------
phylum_df <- psmelt(pseq_phylum_rel)

# ---------------------------------
# 5. Clean phylum labels
# ---------------------------------
  phylum_df <- phylum_df %>%
  dplyr::mutate(
    Phylum = as.character(Phylum),
    Phylum = dplyr::case_when(
      is.na(Phylum) ~ "Unclassified",
      Phylum == "" ~ "Unclassified",
      tolower(Phylum) %in% c("uncultured", "unclassified", "ambiguous_taxa", "unknown") ~ "Unclassified",
      Phylum == "Firmicutes" ~ "Bacillota",
      Phylum %in% c("Actinobacteria", "Actinobacteriota") ~ "Actinomycetota",
      Phylum == "Proteobacteria" ~ "Pseudomonadota",
      TRUE ~ Phylum
    )
  )

phylum_summary <- phylum_df %>%
  dplyr::group_by(Phylum) %>%
  dplyr::summarise(mean_abundance = mean(Abundance, na.rm = TRUE), .groups = "drop")

keep_phyla <- phylum_summary %>%
  dplyr::filter(mean_abundance >= 0.01) %>%
  dplyr::pull(Phylum)

phylum_panel_df <- phylum_df %>%
  dplyr::filter(Phylum %in% keep_phyla)

# ---------------------------------
# 6. Choose top phyla by mean abundance
# ---------------------------------
phylum_summary <- phylum_df %>%
  dplyr::group_by(Phylum) %>%
  dplyr::summarise(mean_abundance = mean(Abundance, na.rm = TRUE), .groups = "drop")

# keep only phyla above a threshold
keep_phyla <- phylum_summary %>%
  dplyr::filter(mean_abundance >= 0.05) %>%   # Lyanna said this should be the threshold 
  dplyr::pull(Phylum)

# filter the plotting dataframe
phylum_panel_df <- phylum_df %>%
  dplyr::filter(Phylum %in% keep_phyla)

# Keep sample order grouped by experimental group
phylum_panel_df <- phylum_df %>%
  mutate(Sample = factor(Sample, levels = unique(Sample[order(experimental_group, Sample)])))

#Summary data (mean and SE)
phylum_summary_stats <- phylum_df %>%
  dplyr::group_by(Phylum, experimental_group) %>%
  dplyr::summarise(
    mean_abundance = mean(Abundance, na.rm = TRUE),
    se_abundance = sd(Abundance, na.rm = TRUE) / sqrt(dplyr::n()),
    .groups = "drop"
  )
#Keep desired phyla 
phylum_summary_stats <- phylum_summary_stats %>%
  filter(Phylum %in% keep_phyla)

#Top Phylum that was not close to 0% relative abundance 
focus_phyla <- c("Bacillota", "Bacteroidota", "Actinomycetota")

phylum_summary_focus <- phylum_summary_stats %>%
  dplyr::filter(Phylum %in% focus_phyla)

# only needed if you later decide to add raw sample points back
phylum_panel_focus <- phylum_panel_df %>%
  dplyr::filter(Phylum %in% focus_phyla)

#Making figure with error bars 
phylum_focus_plot <- ggplot(
  phylum_summary_focus,
  aes(x = experimental_group, y = mean_abundance, fill = experimental_group)
) +
  geom_col(width = 0.6, color = "black") +
  geom_errorbar(
    aes(
      ymin = pmax(mean_abundance - se_abundance, 0),
      ymax = mean_abundance + se_abundance
    ),
    width = 0.2,
    linewidth = 0.6
  ) +
  geom_jitter(
    data = phylum_panel_focus,
    aes(x = experimental_group, y = Abundance),
    width = 0.12,
    size = 1.8,
    alpha = 0.7,
    color = "black",
    inherit.aes = FALSE
  ) +
  facet_wrap(~ Phylum, scales = "fixed", ncol = 2) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(
    values = c(
      "BBS/HC" = "#B3E1F0",
      "BBS/IS" = "#148CBA",
      "M. aurum/HC" = "#98E0B1",
      "M. aurum/IS" = "#21A951"
    ), 
    labels = c(
      "BBS/HC",
      "BBS/IS",
      expression(italic("M. aurum")~ "/HC"),
      expression(italic("M. aurum")~ "/IS")
    ),
    name = "Group"
  ) +
  guides(fill = guide_legend(label.parse = TRUE)) +

  labs(
    x = NULL,
    y = "Relative abundance"
  ) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(angle = 25, hjust = 1),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11)
  )

phylum_focus_plot

ggsave(
  filename = "R_figures/phylum_panels.svg",
  plot = phylum_focus_plot,
  device = svglite::svglite,
  width = 12,
  height = 8,
  units = "in"
)
```
## Relative abundance Treatment: Genus 
```{r}
# -----------------------------
# 1. Agglomerate to Genus
# -----------------------------
pseq_genus <- tax_glom(ps_asv, taxrank = "Genus", NArm = FALSE)

pseq_genus_rel <- transform_sample_counts(
  pseq_genus,
  function(x) x / sum(x)
)

genus_df <- psmelt(pseq_genus_rel)

# -----------------------------
# 2. Clean experimental group labels
# -----------------------------
genus_df$experimental_group <- paste0(
  tolower(as.character(genus_df$Treatment)),
  ".",
  as.character(genus_df$Condition)
)

genus_df$experimental_group <- factor(
  genus_df$experimental_group,
  levels = c("bbs.HC", "bbs.IS", "maurum.HC", "maurum.IS"),
  labels = c("BBS/HC", "BBS/IS", "M. aurum/HC", "M. aurum/IS")
)

# -----------------------------
# 3. Clean genus names
# -----------------------------
genus_df <- genus_df %>%
  dplyr::mutate(
    Genus = as.character(Genus),
    Genus = dplyr::case_when(
      is.na(Genus) ~ "Unclassified",
      Genus == "" ~ "Unclassified",
      tolower(Genus) %in% c(
        "uncultured", "unclassified", "ambiguous_taxa",
        "gut_metagenome", "metagenome", "unknown",
        "na", "n/a"
      ) ~ "Unclassified",
      TRUE ~ Genus
    ),
    Genus = gsub("\\[|\\]", "", Genus)
  )

# -----------------------------
# 4. Select focus genera
# -----------------------------
genus_summary <- genus_df %>%
  dplyr::group_by(Genus) %>%
  dplyr::summarise(
    mean_abundance = mean(Abundance, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(desc(mean_abundance))

head(genus_summary, 20)

focus_genera <- genus_summary %>%
  dplyr::filter(Genus != "Unclassified") %>%
  dplyr::slice_head(n = 12) %>%
  dplyr::pull(Genus)

# keep only genus above a threshold
keep_genus <- genus_summary %>%
  dplyr::filter(mean_abundance >= 0.05) %>%   # Lyanna said this should be the threshold 
  dplyr::pull(Genus)
genus_panel_focus <- genus_df %>%
  dplyr::filter(Genus %in% keep_genus)

# -----------------------------
# 5. Filter to focus genera
# -----------------------------

genus_summary_focus <- genus_panel_focus %>%
  dplyr::group_by(Genus, experimental_group) %>%
  dplyr::summarise(
    n = sum(!is.na(Abundance)),
    mean_abundance = mean(Abundance, na.rm = TRUE),
    sd_abundance = sd(Abundance, na.rm = TRUE),
    se_abundance = sd_abundance / sqrt(n),
    .groups = "drop"
  )

# -----------------------------
# 6. Plot genus relative abundance
# -----------------------------
genus_focus_plot <- ggplot(
  genus_summary_focus,
  aes(x = experimental_group, y = mean_abundance, fill = experimental_group)
) +
  geom_col(width = 0.6, color = "black") +
  geom_errorbar(
    aes(
      ymin = pmax(mean_abundance - se_abundance, 0),
      ymax = mean_abundance + se_abundance
    ),
    width = 0.2,
    linewidth = 0.6
  ) +
  geom_jitter(
    data = genus_panel_focus,
    aes(x = experimental_group, y = Abundance),
    width = 0.12,
    size = 1.8,
    alpha = 0.7,
    color = "black",
    inherit.aes = FALSE
  ) +
  facet_wrap(~ Genus, scales = "fixed", ncol = 3) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.03))
  ) +
  scale_fill_manual(
    values = c(
      "BBS/HC" = "#B3E1F0",
      "BBS/IS" = "#148CBA",
      "M. aurum/HC" = "#98E0B1",
      "M. aurum/IS" = "#21A951"
    ),
    labels = c(
      "BBS/HC",
      "BBS/IS",
      expression(italic("M. aurum")~ "/HC"),
      expression(italic("M. aurum")~ "/IS")
    ),
    name = "Group"
  ) +
  guides(fill = guide_legend(label.parse = TRUE)) +
  labs(
    x = NULL,
    y = "Relative abundance"
  ) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    strip.text = element_text(size = 12, face = "italic"),
    axis.text.x = element_text(angle = 25, hjust = 1, size = 9),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11),
    legend.key.size = unit(0.6, "cm")
  )

genus_focus_plot

ggsave(
  filename = "R_figures/relative_abundance_genus_panels.svg",
  plot = genus_focus_plot,
  device = svglite::svglite,
  width = 12,
  height = 8,
  units = "in"
)
```
## Relative abundance Condition: Phylum
```{r}
# Set condition colors
condition_colors <- c(
  "HC" = "#B3E1F0",
  "IS" = "#148CBA"
)

# Phylum relative abundance
ps_phylum_rel <- transform_sample_counts(ps_phylum, function(x) x / sum(x))
phylum_condition_df <- psmelt(ps_phylum_rel)

# Clean Condition
phylum_condition_df$Condition <- factor(
  as.character(phylum_condition_df$Condition),
  levels = c("HC", "IS")
)

# Update phylum names
phylum_condition_df <- phylum_condition_df %>%
  dplyr::mutate(
    Phylum = as.character(Phylum),
    Phylum = dplyr::case_when(
      Phylum == "Firmicutes" ~ "Bacillota",
      Phylum %in% c("Bacteroidetes", "Bacteroidota") ~ "Bacteroidota",
      Phylum == "Proteobacteria" ~ "Pseudomonadota",
      Phylum %in% c("Actinobacteria", "Actinobacteriota") ~ "Actinomycetota",
      Phylum == "Verrucomicrobia" ~ "Verrucomicrobiota",
      TRUE ~ Phylum
    )
  )

# Focus on dominant phyla
focus_phyla <- c("Bacillota", "Bacteroidota", "Actinomycetota")

phylum_condition_focus <- phylum_condition_df %>%
  dplyr::filter(Phylum %in% focus_phyla)

phylum_condition_summary <- phylum_condition_focus %>%
  dplyr::group_by(Phylum, Condition) %>%
  dplyr::summarise(
    n = sum(!is.na(Abundance)),
    mean_abundance = mean(Abundance, na.rm = TRUE),
    se_abundance = sd(Abundance, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

phylum_condition_plot <- ggplot(
  phylum_condition_summary,
  aes(x = Condition, y = mean_abundance, fill = Condition)
) +
  geom_col(width = 0.6, color = "black") +
  geom_errorbar(
    aes(
      ymin = pmax(mean_abundance - se_abundance, 0),
      ymax = mean_abundance + se_abundance
    ),
    width = 0.2,
    linewidth = 0.6
  ) +
  geom_jitter(
    data = phylum_condition_focus,
    aes(x = Condition, y = Abundance),
    width = 0.12,
    size = 1.8,
    alpha = 0.7,
    color = "black",
    inherit.aes = FALSE
  ) +
  facet_wrap(~ Phylum, scales = "fixed", ncol = 2) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.03))
  ) +
  scale_fill_manual(values = condition_colors, name = "Condition") +
  labs(x = NULL, y = "Relative abundance") +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    strip.text = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11)
  )

phylum_condition_plot

ggsave(
  filename = "R_figures/relative_abundance_phylum_condition_panels.svg",
  plot = phylum_condition_plot,
  device = svglite::svglite,
  width = 12,
  height = 8,
  units = "in"
)

```
## Relative abundance Condition: Genus ALL
```{r}
# Genus relative abundance
ps_genus_rel <- transform_sample_counts(ps_genus, function(x) x / sum(x))
genus_condition_df <- psmelt(ps_genus_rel)

# Clean Condition
genus_condition_df$Condition <- factor(
  as.character(genus_condition_df$Condition),
  levels = c("HC", "IS")
)

# Clean genus names
genus_condition_df <- genus_condition_df %>%
  dplyr::mutate(
    Genus = as.character(Genus),
    Genus = dplyr::case_when(
      is.na(Genus) ~ "Unclassified",
      Genus == "" ~ "Unclassified",
      tolower(Genus) %in% c(
        "uncultured", "unclassified", "ambiguous_taxa",
        "gut_metagenome", "metagenome", "unknown",
        "na", "n/a"
      ) ~ "Unclassified",
      TRUE ~ Genus
    ),
    Genus = gsub("\\[|\\]", "", Genus)
  )

# Summarize mean ± SE
genus_condition_summary <- genus_condition_df %>%
  dplyr::group_by(Genus, Condition) %>%
  dplyr::summarise(
    n = sum(!is.na(Abundance)),
    mean_abundance = mean(Abundance, na.rm = TRUE),
    se_abundance = sd(Abundance, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

genus_condition_plot <- ggplot(
  genus_condition_summary,
  aes(x = Condition, y = mean_abundance, fill = Condition)
) +
  geom_col(width = 0.6, color = "black") +
  geom_errorbar(
    aes(
      ymin = pmax(mean_abundance - se_abundance, 0),
      ymax = mean_abundance + se_abundance
    ),
    width = 0.2,
    linewidth = 0.6
  ) +
  geom_jitter(
    data = genus_condition_df,
    aes(x = Condition, y = Abundance),
    width = 0.12,
    size = 2,
    alpha = 0.75,
    color = "black",
    inherit.aes = FALSE
  ) +
  facet_wrap(~ Genus, scales = "fixed") +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_fill_manual(values = condition_colors, name = "Condition") +
  labs(x = NULL, y = "Relative abundance") +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    strip.text = element_text(size = 12, face = "italic"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11)
  )

genus_condition_plot

ggsave(
  filename = "R_figures/relative_abundance_genus_condition_panels.svg",
  plot = genus_condition_plot,
  device = svglite::svglite,
  width = 12,
  height = 8,
  units = "in"
)
```
## Relative abundance Condition: Genus, targeting significant genus from AMCOM-BC2
```{r}
target_genus <- "Faecalibaculum"

# Genus relative abundance
ps_genus_rel <- transform_sample_counts(ps_genus, function(x) x / sum(x))
genus_condition_df <- psmelt(ps_genus_rel)

# Clean Condition
genus_condition_df$Condition <- factor(
  as.character(genus_condition_df$Condition),
  levels = c("HC", "IS")
)

# Clean genus names
genus_condition_df <- genus_condition_df %>%
  dplyr::mutate(
    Genus = as.character(Genus),
    Genus = dplyr::case_when(
      is.na(Genus) ~ "Unclassified",
      Genus == "" ~ "Unclassified",
      tolower(Genus) %in% c(
        "uncultured", "unclassified", "ambiguous_taxa",
        "gut_metagenome", "metagenome", "unknown",
        "na", "n/a"
      ) ~ "Unclassified",
      TRUE ~ Genus
    ),
    Genus = gsub("\\[|\\]", "", Genus)
  )

# Filter to significant genus
genus_condition_focus <- genus_condition_df %>%
  dplyr::filter(Genus == target_genus)

# Summarize mean ± SE
genus_condition_summary <- genus_condition_focus %>%
  dplyr::group_by(Genus, Condition) %>%
  dplyr::summarise(
    n = sum(!is.na(Abundance)),
    mean_abundance = mean(Abundance, na.rm = TRUE),
    se_abundance = sd(Abundance, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

genus_condition_focus_plot <- ggplot(
  genus_condition_summary,
  aes(x = Condition, y = mean_abundance, fill = Condition)
) +
  geom_col(width = 0.6, color = "black") +
  geom_errorbar(
    aes(
      ymin = pmax(mean_abundance - se_abundance, 0),
      ymax = mean_abundance + se_abundance
    ),
    width = 0.2,
    linewidth = 0.6
  ) +
  geom_jitter(
    data = genus_condition_focus,
    aes(x = Condition, y = Abundance),
    width = 0.12,
    size = 2,
    alpha = 0.75,
    color = "black",
    inherit.aes = FALSE
  ) +
  facet_wrap(~ Genus, scales = "fixed") +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_fill_manual(values = condition_colors, name = "Condition") +
  labs(x = NULL, y = "Relative abundance") +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black"),
    strip.text = element_text(size = 12, face = "italic"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11)
  )

genus_condition_focus_plot

ggsave(
  filename = "R_figures/relative_abundance_genus_target_condition_panels.svg",
  plot = genus_condition_plot,
  device = svglite::svglite,
  width = 12,
  height = 8,
  units = "in"
)
```
## Statistical Analysis: ANCOM-BC2 Treatment and Condition tables  
```{r}
# -----------------------------
# 1. Make sure factors are set
# -----------------------------
sample_data(ps_asv)$Condition <- factor(
  as.character(sample_data(ps_asv)$Condition),
  levels = c("HC", "IS")
)

sample_data(ps_asv)$Treatment <- trimws(as.character(sample_data(ps_asv)$Treatment))
sample_data(ps_asv)$Treatment[sample_data(ps_asv)$Treatment == "BBS"] <- "bbs"
sample_data(ps_asv)$Treatment[sample_data(ps_asv)$Treatment == "M. aurum"] <- "maurum"

sample_data(ps_asv)$Treatment <- factor(
  sample_data(ps_asv)$Treatment,
  levels = c("bbs", "maurum")
)

# -----------------------------
# 2. Aggregate to phylum + genus
# -----------------------------
ps_phylum <- tax_glom(ps_asv, taxrank = "Phylum", NArm = FALSE)
ps_genus  <- tax_glom(ps_asv, taxrank = "Genus",  NArm = FALSE)

# -----------------------------
# 3. ANCOM-BC2 runner
# -----------------------------
run_ancombc2_level <- function(ps_obj, fix_formula, group_var) {
  ancombc2(
    data = ps_obj,
    tax_level = NULL,
    fix_formula = fix_formula,
    rand_formula = NULL,
    group = group_var,
    p_adj_method = "BH",   # FDR correction
    prv_cut = 0.10,
    lib_cut = 1000,
    struc_zero = TRUE,
    neg_lb = TRUE,
    pseudo_sens = TRUE,
    alpha = 0.05,
    n_cl = 1,
    verbose = TRUE
  )
}

# -----------------------------
# 4. Run ANCOM-BC2
# -----------------------------
ancom2_phylum_condition <- run_ancombc2_level(ps_phylum, "Condition", "Condition")
ancom2_genus_condition  <- run_ancombc2_level(ps_genus,  "Condition", "Condition")

ancom2_phylum_treatment <- run_ancombc2_level(ps_phylum, "Treatment", "Treatment")
ancom2_genus_treatment  <- run_ancombc2_level(ps_genus,  "Treatment", "Treatment")

# --------------------------
# 5. Extraction function 
# --------------------------
extract_ancombc2_table <- function(ancom_obj, ps_obj, rank_name, effect_name,
                                   comparison_label, reference_label, comparison_group_label) {

  res_df <- as.data.frame(ancom_obj$res)

  tax_df <- as.data.frame(phyloseq::tax_table(ps_obj)) %>%
    tibble::rownames_to_column("TaxonID")

  lfc_col  <- paste0("lfc_", effect_name)
  se_col   <- paste0("se_", effect_name)
  stat_col <- paste0("W_", effect_name)
  p_col    <- paste0("p_", effect_name)
  q_col    <- paste0("q_", effect_name)
  diff_col <- paste0("diff_", effect_name)
  ss_col   <- paste0("passed_ss_", effect_name)

  out <- res_df %>%
    dplyr::rename(TaxonID = taxon) %>%
    dplyr::left_join(tax_df, by = "TaxonID") %>%
    dplyr::mutate(
      taxonomic_level = rank_name,
      taxon_name = as.character(.data[[rank_name]]),
      comparison = comparison_label,
      reference_group = reference_label,
      comparison_group = comparison_group_label,
      log_fold_change = .data[[lfc_col]],
      standard_error = .data[[se_col]],
      test_statistic = .data[[stat_col]],
      p_value = .data[[p_col]],
      fdr_q_value = .data[[q_col]],
      differentially_abundant = .data[[diff_col]]
    )

  # apply phylum name updates outside mutate
  if (rank_name == "Phylum") {
    out$taxon_name <- update_phylum_names(out$taxon_name)
  }

  # add optional sensitivity column safely
  if (ss_col %in% colnames(out)) {
    out$passed_sensitivity <- out[[ss_col]]
  } else {
    out$passed_sensitivity <- NA
  }

  out <- out %>%
    dplyr::mutate(
      significant_fdr_0.05 = dplyr::if_else(
        !is.na(fdr_q_value) & fdr_q_value < 0.05,
        "Yes", "No"
      ),
      direction = dplyr::case_when(
        log_fold_change > 0 ~ paste0("Higher in ", comparison_group_label),
        log_fold_change < 0 ~ paste0("Higher in ", reference_label),
        TRUE ~ "No change"
      )
    ) %>%
    dplyr::filter(
      !is.na(taxon_name),
      taxon_name != "",
      !tolower(taxon_name) %in% c(
        "na", "n/a", "uncultured", "unclassified",
        "ambiguous_taxa", "gut_metagenome", "metagenome",
        "unknown", "uncultured_bacterium", "uncultured_organism"
      )
    ) %>%
    dplyr::mutate(
      taxon_name = gsub("\\[|\\]", "", taxon_name),
      log_fold_change = round(log_fold_change, 3),
      standard_error = round(standard_error, 3),
      test_statistic = round(test_statistic, 3),
      p_value = signif(p_value, 3),
      fdr_q_value = signif(fdr_q_value, 3)
    ) %>%
    dplyr::select(
      taxonomic_level,
      taxon_name,
      comparison,
      reference_group,
      comparison_group,
      log_fold_change,
      standard_error,
      test_statistic,
      p_value,
      fdr_q_value,
      significant_fdr_0.05,
      differentially_abundant,
      passed_sensitivity,
      direction,
      TaxonID
    ) %>%
    dplyr::arrange(fdr_q_value, dplyr::desc(abs(log_fold_change)))

  out
}
#updating phylum names 
update_phylum_names <- function(x) {
  dplyr::case_when(
    x == "Firmicutes" ~ "Bacillota",
    x %in% c("Bacteroidetes", "Bacteroidota") ~ "Bacteroidota",
    x == "Proteobacteria" ~ "Pseudomonadota",
    x %in% c("Actinobacteria", "Actinobacteriota") ~ "Actinomycetota",
    x == "Fusobacteria" ~ "Fusobacteriota",
    x == "Lentisphaerae" ~ "Lentisphaerota",
    x == "Tenericutes" ~ "Mycoplasmatota",
    x == "Synergistetes" ~ "Synergistota",
    x == "Verrucomicrobia" ~ "Verrucomicrobiota",
    TRUE ~ x
  )
}                                  
# ----------------------
# 6. Making tables
# ----------------------
phylum_condition_table <- extract_ancombc2_table(
  ancom_obj = ancom2_phylum_condition,
  ps_obj = ps_phylum,
  rank_name = "Phylum",
  effect_name = "ConditionIS",
  comparison_label = "IS vs HC",
  reference_label = "HC",
  comparison_group_label = "IS"
)

genus_condition_table <- extract_ancombc2_table(
  ancom_obj = ancom2_genus_condition,
  ps_obj = ps_genus,
  rank_name = "Genus",
  effect_name = "ConditionIS",
  comparison_label = "IS vs HC",
  reference_label = "HC",
  comparison_group_label = "IS"
)

phylum_treatment_table <- extract_ancombc2_table(
  ancom_obj = ancom2_phylum_treatment,
  ps_obj = ps_phylum,
  rank_name = "Phylum",
  effect_name = "Treatmentmaurum",
  comparison_label = "M. aurum vs BBS",
  reference_label = "BBS",
  comparison_group_label = "M. aurum"
)

genus_treatment_table <- extract_ancombc2_table(
  ancom_obj = ancom2_genus_treatment,
  ps_obj = ps_genus,
  rank_name = "Genus",
  effect_name = "Treatmentmaurum",
  comparison_label = "M. aurum vs BBS",
  reference_label = "BBS",
  comparison_group_label = "M. aurum"
)
ancombc2_publication_table <- dplyr::bind_rows(
  phylum_condition_table,
  genus_condition_table,
  phylum_treatment_table,
  genus_treatment_table
)

# ---------------------------
# 7. Exporting tables to Excel
# ---------------------------

dir.create("R_tables", showWarnings = FALSE)

wb <- openxlsx::createWorkbook()

openxlsx::addWorksheet(wb, "Phylum_Condition")
openxlsx::writeData(wb, "Phylum_Condition", phylum_condition_table)

openxlsx::addWorksheet(wb, "Genus_Condition")
openxlsx::writeData(wb, "Genus_Condition", genus_condition_table)

openxlsx::addWorksheet(wb, "Phylum_Treatment")
openxlsx::writeData(wb, "Phylum_Treatment", phylum_treatment_table)

openxlsx::addWorksheet(wb, "Genus_Treatment")
openxlsx::writeData(wb, "Genus_Treatment", genus_treatment_table)

openxlsx::addWorksheet(wb, "Combined")
openxlsx::writeData(wb, "Combined", ancombc2_publication_table)

# Optional formatting
header_style <- openxlsx::createStyle(
  textDecoration = "bold",
  fgFill = "#D9EAF7",
  border = "Bottom"
)

for (sheet in names(wb)) {
  openxlsx::addStyle(
    wb, sheet,
    style = header_style,
    rows = 1,
    cols = 1:ncol(openxlsx::readWorkbook(wb, sheet = sheet)),
    gridExpand = TRUE
  )
  openxlsx::freezePane(wb, sheet, firstRow = TRUE)
  openxlsx::setColWidths(wb, sheet, cols = 1:20, widths = "auto")
}

openxlsx::saveWorkbook(
  wb,
  file = "R_tables/ANCOMBC2_phylum_genus_publication_tables.xlsx",
  overwrite = TRUE
)
```
## Using ANCOM-BC tables to make figure: Treatment, higher and lower on the Phylum and Genus levels
```{r}
# Not really prefered since it can be messy and intrepretation may confuse readers. Provided it since I presented this data for lab meeting

# Keep internal coding stable for analysis
sample_data(ps_asv)$experimental_group <- factor(
  as.character(sample_data(ps_asv)$experimental_group),
  levels = c("bbs.HC", "bbs.IS", "maurum.HC", "maurum.IS")
)

sample_data(ps_asv)$Condition <- factor(
  as.character(sample_data(ps_asv)$Condition),
  levels = c("HC", "IS")
)

# aggregate to phylum and genus
ps_phylum <- tax_glom(ps_asv, taxrank = "Phylum", NArm = FALSE)
ps_genus  <- tax_glom(ps_asv, taxrank = "Genus",  NArm = FALSE)

# BBS-only subset for the condition-only comparison
ps_asv_bbs <- subset_samples(ps_asv, Treatment %in% c("bbs", "BBS"))
ps_phylum_bbs <- tax_glom(ps_asv_bbs, taxrank = "Phylum", NArm = FALSE)
ps_genus_bbs  <- tax_glom(ps_asv_bbs, taxrank = "Genus",  NArm = FALSE)

sample_data(ps_asv_bbs)$Condition <- factor(
  as.character(sample_data(ps_asv_bbs)$Condition),
  levels = c("HC", "IS")
)

#ANCOM-BC runner and main framework
run_ancom_level <- function(ps_obj, formula_var, group_var) {
  ancombc(
    data = ps_obj,
    tax_level = NULL,
    formula = formula_var,
    p_adj_method = "BH",   # this gives FDR-adjusted q-values
    prv_cut = 0.10,
    lib_cut = 1000,
    group = group_var,
    struc_zero = TRUE,
    neg_lb = TRUE,
    tol = 1e-5,
    max_iter = 100,
    conserve = TRUE,
    alpha = 0.05,
    global = FALSE
  )
}

#Experimental group model and reference gorup will be bbs.hc
ancom_phylum_groups <- run_ancom_level(ps_phylum, "experimental_group", "experimental_group")
ancom_genus_groups  <- run_ancom_level(ps_genus,  "experimental_group", "experimental_group")

#Extracting ANCOM-BC Results
extract_ancom_results <- function(ancom_obj, effect_col, ps_obj, rank_name) {
  lfc_df <- as.data.frame(ancom_obj$res$lfc) %>%
    dplyr::rename(TaxonID = taxon)

  p_df <- as.data.frame(ancom_obj$res$p_val) %>%
    dplyr::rename(TaxonID = taxon)

  q_df <- as.data.frame(ancom_obj$res$q_val) %>%
    dplyr::rename(TaxonID = taxon)

  da_df <- as.data.frame(ancom_obj$res$diff_abn) %>%
    dplyr::rename(TaxonID = taxon)

  tax_df <- as.data.frame(tax_table(ps_obj)) %>%
    tibble::rownames_to_column("TaxonID")

  out <- lfc_df %>%
    dplyr::select(TaxonID, lfc = all_of(effect_col)) %>%
    dplyr::left_join(
      p_df %>% dplyr::select(TaxonID, p_val = all_of(effect_col)),
      by = "TaxonID"
    ) %>%
    dplyr::left_join(
      q_df %>% dplyr::select(TaxonID, q_val = all_of(effect_col)),
      by = "TaxonID"
    ) %>%
    dplyr::left_join(
      da_df %>% dplyr::select(TaxonID, diff_abn = all_of(effect_col)),
      by = "TaxonID"
    ) %>%
    dplyr::left_join(tax_df, by = "TaxonID")

  if (rank_name %in% colnames(out)) {
    out$Label <- as.character(out[[rank_name]])
  } else {
    out$Label <- out$TaxonID
  }

  out
}

#Clean taxa and remove "junk labels"
clean_taxa_all <- function(df) {
  bad_vals <- c(
    "na", "n/a", "uncultured", "unclassified",
    "ambiguous_taxa", "gut_metagenome", "metagenome",
    "unknown", "uncultured_bacterium", "uncultured_organism"
  )

  df %>%
    mutate(
      Label = trimws(as.character(Label)),
      Label_lower = tolower(Label)
    ) %>%
    filter(
      !is.na(Label),
      Label != "",
      !Label_lower %in% bad_vals
    ) %>%
    mutate(
      FDR_sig = ifelse(!is.na(q_val) & q_val < 0.05, "FDR < 0.05", "NS"),
      Direction = ifelse(lfc > 0, "Positive", "Negative")
    ) %>%
    select(-Label_lower)
}

#Build phylum and genus result tables for the treatment group comparisons 

#Phylum
phylum_bbsIS_vs_bbsHC <- extract_ancom_results(
  ancom_phylum_groups, "experimental_groupbbs.IS", ps_phylum, "Phylum"
) %>% clean_taxa_all()

phylum_maurumHC_vs_bbsHC <- extract_ancom_results(
  ancom_phylum_groups, "experimental_groupmaurum.HC", ps_phylum, "Phylum"
) %>% clean_taxa_all()

phylum_maurumIS_vs_bbsHC <- extract_ancom_results(
  ancom_phylum_groups, "experimental_groupmaurum.IS", ps_phylum, "Phylum"
) %>% clean_taxa_all()

#Genus
genus_bbsIS_vs_bbsHC <- extract_ancom_results(
  ancom_genus_groups, "experimental_groupbbs.IS", ps_genus, "Genus"
) %>% clean_taxa_all()

genus_maurumHC_vs_bbsHC <- extract_ancom_results(
  ancom_genus_groups, "experimental_groupmaurum.HC", ps_genus, "Genus"
) %>% clean_taxa_all()

genus_maurumIS_vs_bbsHC <- extract_ancom_results(
  ancom_genus_groups, "experimental_groupmaurum.IS", ps_genus, "Genus"
) %>% clean_taxa_all()

#Clean up with no brackets for genus labels 
genus_bbsIS_vs_bbsHC$Label <- gsub("\\[|\\]", "", genus_bbsIS_vs_bbsHC$Label)
genus_maurumHC_vs_bbsHC$Label <- gsub("\\[|\\]", "", genus_maurumHC_vs_bbsHC$Label)
genus_maurumIS_vs_bbsHC$Label <- gsub("\\[|\\]", "", genus_maurumIS_vs_bbsHC$Label)


#Plot all taxa and marks which ones pass FDR
plot_ancom_all <- function(df, title_text, y_label, neg_label, pos_label) {
  plot_df <- df %>%
    filter(!is.na(lfc), !is.na(Label)) %>%
    arrange(lfc)

  plot_df$Label <- factor(plot_df$Label, levels = plot_df$Label)

  ggplot(plot_df, aes(x = lfc, y = Label, fill = lfc > 0)) +
    geom_col(width = 0.9, color = "black") +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_point(
      data = subset(plot_df, q_val < 0.05),
      aes(x = lfc, y = Label),
      inherit.aes = FALSE,
      size = 2.2,
      shape = 21,
      fill = "black",
      color = "black"
    ) +
    scale_fill_manual(
      values = c("TRUE" = "#D55E00", "FALSE" = "#0072B2"),
      name = "Direction",
      labels = c("FALSE" = neg_label, "TRUE" = pos_label)
    ) +
    theme_bw() +
    labs(
      x = "Log Fold Change",
      y = y_label,
      title = title_text
    ) +
    theme(
      axis.text.y = element_text(size = 8, face = "italic"),
      axis.text.x = element_text(size = 11),
      axis.title = element_text(size = 12),
      plot.title = element_text(size = 13, face = "bold"),
      legend.position = "top",
      plot.margin = margin(10, 20, 10, 20)
    )
}

#Making figures for Phylum
p_phylum_1 <- plot_ancom_all(
  phylum_bbsIS_vs_bbsHC,
  "BBS/IS vs BBS/HC",
  "Phylum",
  "Higher in BBS/HC",
  "Higher in BBS/IS"
)

p_phylum_2 <- plot_ancom_all(
  phylum_maurumHC_vs_bbsHC,
  "M. aurum/HC vs BBS/HC",
  "Phylum",
  "Higher in BBS/HC",
  "Higher in M. aurum/HC"
)

p_phylum_3 <- plot_ancom_all(
  phylum_maurumIS_vs_bbsHC,
  "M. aurum/IS vs BBS/HC",
  "Phylum",
  "Higher in BBS/HC",
  "Higher in M. aurum/IS"
)


figure_phylum <- (p_phylum_1 | p_phylum_2) / (p_phylum_3 )
figure_phylum

#Making figures for genus 
p_genus_1 <- plot_ancom_all(
  genus_bbsIS_vs_bbsHC,
  "BBS/IS vs BBS/HC",
  "Genus",
  "Higher in BBS/HC",
  "Higher in BBS/IS"
)

p_genus_2 <- plot_ancom_all(
  genus_maurumHC_vs_bbsHC,
  "M. aurum/HC vs BBS/HC",
  "Genus",
  "Higher in BBS/HC",
  "Higher in M. aurum/HC"
)

p_genus_3 <- plot_ancom_all(
  genus_maurumIS_vs_bbsHC,
  "M. aurum/IS vs BBS/HC",
  "Genus",
  "Higher in BBS/HC",
  "Higher in M. aurum/IS"
)


figure_genus <- (p_genus_1 | p_genus_2) / (p_genus_3)
figure_genus

ggsave(
  filename = "R_figures/ancombc_phylum_panels.svg",
  plot = figure_phylum,
  device = svglite::svglite,
  width = 14,
  height = 10,
  units = "in"
)

ggsave(
  filename = "R_figures/ancombc_genus_panels.svg",
  plot = figure_genus,
  device = svglite::svglite,
  width = 14,
  height = 10,
  units = "in"
)
```
## Using ANCOM-BC tables to make figures: Condition, higher and lower on the Phylum and Genus levels
```{r}

run_ancom_level <- function(ps_obj, formula_var, group_var) {
  ancombc(
    data = ps_obj,
    tax_level = NULL,
    formula = formula_var,
    p_adj_method = "BH",   # this gives FDR-adjusted q-values
    prv_cut = 0.10,
    lib_cut = 1000,
    group = group_var,
    struc_zero = TRUE,
    neg_lb = TRUE,
    tol = 1e-5,
    max_iter = 100,
    conserve = TRUE,
    alpha = 0.05,
    global = FALSE
  )
}

ancom_phylum_bbs_condition <- run_ancom_level(ps_phylum_bbs, "Condition", "Condition")
ancom_genus_bbs_condition  <- run_ancom_level(ps_genus_bbs,  "Condition", "Condition") 


# Comparsions tables 
phylum_bbs_condition <- extract_ancom_results(
  ancom_phylum_bbs_condition, "ConditionIS", ps_phylum_bbs, "Phylum"
) %>% clean_taxa_all()

genus_bbs_condition <- extract_ancom_results(
  ancom_genus_bbs_condition, "ConditionIS", ps_genus_bbs, "Genus"
) %>% clean_taxa_all()


genus_bbs_condition$Label <- gsub("\\[|\\]", "", genus_bbs_condition$Label)

p_phylum_4 <- plot_ancom_all(
  phylum_bbs_condition,
  "BBS Condition: IS vs HC",
  "Phylum",
  "Higher in BBS/HC",
  "Higher in BBS/IS"
)

p_genus_4 <- plot_ancom_all(
  genus_bbs_condition,
  "BBS Condition: IS vs HC",
  "Genus",
  "Higher in BBS/HC",
  "Higher in BBS/IS"
)

```
